---
title: "Assing CSS to SVG"
author: "ryan.benac"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: inline
---

# Install Packages
```{r setup, eval=FALSE}
remotes::install_github("coolbutuseless/minisvg") # Handles reading and modification of SVG
remotes::install_github("coolbutuseless/minicss") # Handles formatting css
```

```{r}
library(minisvg)
library(minicss)
```

## Import and View the SVG
I removed the fill style from the right rectangle for testing.
```{r}
# Read the SVG as a minisvg SVG Document
svg_panama <- minisvg::parse_svg_doc("man/figures/panama_canal.svg")
minisvg::knit_print.SVGDocument(svg_panama)
```

# Add manually scripted CSS to right rectancle

When creating CSS, be sure to use !important if the SVG or id within the SCG already contain a style definition.

```{r}
# Add manual CSS
svg_panama$add_css("
#comp3 {
  fill: green !important;
}")
minisvg::knit_print.SVGDocument(svg_panama)

# Adding !important overwrites any existing style for that shape (id)
```

```{r}
# use an external style sheet
svg_panama <- minisvg::parse_svg_doc("man/figures/panama_canal.svg")

my_css <- "
#comp3 {
  fill: yellow !important;
}"
writeLines(my_css, "css/external.css")
svg_panama$add_css_url("css/external.css")

# save svg and reload
writeLines(as.character(svg_panama), "man/figures/yellow_panama.svg")
```

Rmd will not show SVG with external CSS definition. Saving and opening "yellow_panama.svg" confirms that this chunk executed correctly.


# Creating CSS with cssparser
Create the CSS, append to add id selector with #, then append with the text version of 
```{r}
# create css by giving it our id (comp3) and the color to fill it with
svg_panama <- minisvg::parse_svg_doc("man/figures/panama_canal.svg")

this_id <- "comp3"
color_def <- '#d60000'
new_css <- Style$new(paste("#", this_id, sep=""), fill = color_def)


# add that css to the svg
svg_panama$add_css(as.character(new_css))
minisvg::knit_print.SVGDocument(svg_panama)
```

# Adding animations with minicss
```{r}
svg_panama <- minisvg::parse_svg_doc("man/figures/panama_canal.svg")

# first step, then modify to add move
key_name = "color_change"
my_frame <- minicss::Keyframes$new(name = key_name)

my_frame$add(time = "0%", fill = 'blue')
my_frame$add(time = "25%", fill = 'orange')
my_frame$add(time = "50%", fill = 'yellow')
my_frame$add(time = "75%", fill = 'green')
my_frame$add(time = "100%", fill = 'blue')

##### add that keyframe declaration into the svg
# create stylesheet
#stylesheet <- minicss::css_stylesheet()
svg_panama$add_css(as.character(my_frame))

print(svg_panama)
```
Let's add an animation declaration
```{r}
###### Add animation declaration
this_id <- "comp3"
anim_dur <- 2
iter_count = 'infinite'
dir = 'alternate'

# define the css animation
this_id <- "comp3"
new_css <- Style$new(paste("#", this_id, sep=""))

# update with animation declaration
new_css$update(css_prop$animation(name = key_name, duration = anim_dur, iteration_count = iter_count, direction = dir))

# add that new css to the svg
svg_panama$add_css(as.character(new_css))

#minisvg::knit_print.SVGDocument(svg_panama)
print(svg_panama)

# save to view
writeLines(as.character(svg_panama), "man/figures/keyframe_panama.svg")
```
#Potential Issues
1. When using prestyled CSS, this code may need to temporarily overwrite that css definition with an !important declaration. This may be helpful later depending on how the final SVG is formatted. See here for more info <https://www.w3docs.com/snippets/css/how-to-override-css-styles.html>.
2. How will we pull out each id from the svg. Use a naming convention when making the id? That will make it easier to parse through and find what we want. We could then add a description in the id after a separator "_" and build a df in r with important id information.


# Find and extract svg group id
```{r}
# reload fresh svg
svg_panama <- minisvg::parse_svg_doc("man/figures/panama_canal.svg")

# find children by index
svg_panama$children
```
```{r}
# isolate to the drawing definitions
svg_panama$children[[4]]

### note that this might be different depending on the SVG definition. Requires further testing
```
```{r}
# pull only the id from that group
draw_group <- svg_panama$children[[4]]
draw_group <- draw_group$children
for (i in 1:length(draw_group)) {
  print(draw_group[[i]]$attribs$id)
}
```
```{r}
# add those id into a 
# create empty list
id_list <- vector(mode='list', length=length(draw_group))

# iterate then append values
for (i in 1:length(draw_group)) {
  id_list[i] <- draw_group[[i]]$attribs$id
}

id_list
```

# Create a function that takes in svg doc and returns vector of shape id
```{r}
get_ids <- function(SVG_name) {
  draw_group <- SVG_name$children[[4]]$children
  
  # create empty list
  df <- data.frame(matrix(ncol = 0, nrow = length(draw_group)))
  df$id <- ""
  
  # iterate then append values
  for (i in 1:length(draw_group)) {
    df[i,1] <- draw_group[[i]]$attribs$id
  }
  
  return(df)
}

svg_panama_desc <- minisvg::parse_svg_doc("man/figures/panama_canal_descriptive_id.svg")

ids_df <- get_ids(svg_panama_desc)
ids_df
```
