---
title: "Animating SVG with Data"
author: "ryan.benac"
date: "`r Sys.Date()`"
output: html_document
---

## Animating an SVG with Data

# Install Packages
```{r setup, eval=FALSE}
remotes::install_github("coolbutuseless/minisvg") # Handles reading and modification of SVG
remotes::install_github("coolbutuseless/minicss") # Handles formatting css
install.packages("colorspace", repos = "http://R-Forge.R-project.org") # Handles color hughes
```

```{r}
library(minisvg)
library(minicss)
library(colorspace)
```

# Loading the SVG
First, we need to load the SVG with labels
```{r}
# function to pull out shape id from the svg
get_ids <- function(SVG_name) {
  draw_group <- SVG_name$children[[4]]$children
  
  # create empty list
  df <- data.frame(matrix(ncol = 0, nrow = length(draw_group)))
  df$id <- ""
  
  # iterate then append values
  for (i in 1:length(draw_group)) {
    df[i,1] <- draw_group[[i]]$attribs$id
  }
  
  return(df)
}

svg_panama <- minisvg::parse_svg_doc("man/figures/panama_canal_descriptive_id.svg")

df_svg_id <- get_ids(svg_panama)
df_svg_id

```

# Create Test Data for Visualization
```{r}
# create empty data frame
df_size <- 100
sal_max <- 32
sal_min <- 1

df_dummyData <- data.frame(matrix(ncol = 0, nrow = df_size))

# create time sequence with readings over 1 day
df_dummyData$time <- seq(as.POSIXct("2023-3-1 0:00"), by="hour", length = df_size)

# create data that builds up (1-100)
df_dummyData$sal_buildUp <- seq(sal_min,sal_max, length=df_size)

# create normally distributed data
set.seed(1)
df_dummyData$sal_norm <- rnorm(df_size, mean=sal_max/2, sd=5)

# create randomly distributed data
df_dummyData$sal_random <- runif(df_size, min=sal_min, max=sal_max)

head(df_dummyData)
```

Make sure the data frame is sorted in order by time. This makes it easier to iterate through and animate in the correct order.
```{r}
# sort dummy data by time
df_dummyData <- df_dummyData[order(df_dummyData$time),]

# add key frame id to sorted df. This will be the key frame animation id from 1 to df_size

df_dummyData$keyframe_id <- seq(from=1, to=100, length = df_size)
head(df_dummyData)
```

# Applying Dummy Data to CSS
Use the dummy data to create key frames. Iterate through the df (because we need to build the css line by line) and assign color based on value

```{r} 
##############################################   BUILD   #############################################################

# reread svg to reset any previous animations
svg_panama <- minisvg::parse_svg_doc("man/figures/panama_canal_descriptive_id.svg")

# first step, then modify to add move
key_name = "anim_svg_build"
my_frame <- minicss::Keyframes$new(name = key_name)

data_col <- "sal_buildUp"
light_blue <- "#ADD8E6"
dark_blue <- "#00008b"

# iterate through df and assign animation
for (row in 1:nrow(df_dummyData)) {
  inst_value <- df_dummyData[row, data_col] # pull out instrument value at this time
  darkness <- inst_value / sal_max
  colorVal <- colorspace::darken(light_blue, darkness) # darken the color by instrument value
  my_frame$add(time = paste(df_dummyData[row,"keyframe_id"], "%",sep = ""), # this will assign keyframe id (between 1-100) to the animation time
               fill = colorVal)
}


##### add that keyframe declaration into the svg
svg_panama$add_css(as.character(my_frame))

print(svg_panama)
```

Next, we need to add an animation declaration for animating our svg object with the keyframe css.
```{r}
###### Add animation declaration
this_id <- "comp3_rightRectangle"
anim_dur <- 15                      # animation duration in seconds
iter_count = 'infinite'
dir = 'normal'

# define the css animation
new_css <- Style$new(paste("#", this_id, sep=""))

# update with animation declaration
new_css$update(css_prop$animation(name = key_name, duration = anim_dur, iteration_count = iter_count, direction = dir))

# add that new css to the svg
svg_panama$add_css(as.character(new_css))

print(svg_panama)
```

Let's also build animations for the middle and left rectangles using the other sample data. This may need to be condensed into a function.
```{r echo=FALSE}
##############################################   NORMAL   #############################################################

# first step, then modify to add move
key_name = "anim_svg_normal"
my_frame <- minicss::Keyframes$new(name = key_name)

data_col <- "sal_norm"
light_blue <- "#ADD8E6"
dark_blue <- "#00008b"

# iterate through df and assign animation
for (row in 1:nrow(df_dummyData)) {
  inst_value <- df_dummyData[row, data_col] # pull out instrument value at this time
  darkness <- inst_value / sal_max
  colorVal <- colorspace::darken(light_blue, darkness) # darken the color by instrument value
  my_frame$add(time = paste(df_dummyData[row,"keyframe_id"], "%",sep = ""), # this will assign keyframe id (between 1-100) to the animation time
               fill = colorVal)
}

##### add that keyframe declaration into the svg
svg_panama$add_css(as.character(my_frame))

###### Add animation declaration
this_id <- "comp2_middleRectangle"
anim_dur <- 15                      # animation duration in seconds
iter_count = 'infinite'
dir = 'normal'

# define the css animation
new_css <- Style$new(paste("#", this_id, sep=""))

# update with animation declaration
new_css$update(css_prop$animation(name = key_name, duration = anim_dur, iteration_count = iter_count, direction = dir))

# add that new css to the svg
svg_panama$add_css(as.character(new_css))

##############################################   RANDOM   #############################################################
# first step, then modify to add move
key_name = "anim_svg_random"
my_frame <- minicss::Keyframes$new(name = key_name)

data_col <- "sal_random"
light_blue <- "#ADD8E6"
dark_blue <- "#00008b"

# iterate through df and assign animation
for (row in 1:nrow(df_dummyData)) {
  inst_value <- df_dummyData[row, data_col] # pull out instrument value at this time
  darkness <- inst_value / sal_max
  colorVal <- colorspace::darken(light_blue, darkness) # darken the color by instrument value
  my_frame$add(time = paste(df_dummyData[row,"keyframe_id"], "%",sep = ""), # this will assign keyframe id (between 1-100) to the animation time
               fill = colorVal)
}

##### add that keyframe declaration into the svg
svg_panama$add_css(as.character(my_frame))

###### Add animation declaration
this_id <- "comp1_leftRectangle"
anim_dur <- 15                      # animation duration in seconds
iter_count = 'infinite'
dir = 'normal'

# define the css animation
new_css <- Style$new(paste("#", this_id, sep=""))

# update with animation declaration
new_css$update(css_prop$animation(name = key_name, duration = anim_dur, iteration_count = iter_count, direction = dir))

# add that new css to the svg
svg_panama$add_css(as.character(new_css))
```

This animates the image to that the rectangles are animated based on:
1. The right rectangle is a build up, meaning at the beginning of time it has a low value and has an even sequence building to 32.
2. The middle rectangle is normally distributed data.
3. The left rectangle is randomly distributed between 1 and 32.

A light blue signifies low values (low salinity) and as it gets darker, the value is increasing to 32 (high salinity). 

# Change Text to Show Time Change
```{r echo=FALSE}
# first step, then modify to add move
key_name = "anim_svg_text"
my_frame <- minicss::Keyframes$new(name = key_name)

data_col <- "time"

# iterate through df and assign animation
for (row in 1:nrow(df_dummyData)) {
  inst_value <- paste('"', df_dummyData[row, data_col], '"', sep="") # pull out instrument value at this time
  
  my_frame$add(time = paste(df_dummyData[row,"keyframe_id"], "%",sep = ""), # this will assign keyframe id (between 1-100) to the animation time
               content = inst_value)
}

##### add that keyframe declaration into the svg
svg_panama$add_css(as.character(my_frame))

###### Add animation declaration
this_id <- "time_text"
anim_dur <- 15                      # animation duration in seconds
iter_count = 'infinite'
#dir = 'normal'

# define the css animation
new_css <- Style$new(paste("#", this_id, sep=""))

# update with animation declaration
new_css$update(css_prop$animation(name = key_name, duration = anim_dur, iteration_count = iter_count, direction = dir))
#new_css$update(css_prop$content(name = key_name, content = ""))


# add that new css to the svg
svg_panama$add_css(as.character(new_css))
```

# Save the SVG to view it
```{r}
# save to view
writeLines(as.character(svg_panama), "man/figures/keyframe_panama_animated.svg")
```


-------------------------------------------------------------------------------------------------------------------------------------------------------



Let's take it up a notch. We not have roughly cleaned clocking and instrumentation data. Let's import a cross section view of one lock, isolate the data from that time, and color code a keyframe definition to match changes in instrumentation readings over the locking time.

I used Inkscape to create a 3x5 representation of instruments in a lock. There are 3 groupings of instruments each with 5 sensors at various depths. I gave each object box a name with as STATION_INSTRUMENT. For example, at station UC1 Instrument C, the naming convention is UC1_C. I downloaded this file as an SVG.
```{R}
# import as an svg_doc object
svg_upper <- minisvg::parse_svg_doc("man/figures/upper_lock_cross_section.svg")

df_upper_svgId <- get_ids(svg_upper)
head(df_upper_svgId)
```

We'll also need to read in the data.
```{R}
df_sgr <- read.table("C:\\Workspace\\GIT\\DATA\\Infographic\\df_sgr.csv")
head(df_sgr)
```



```{R}
unique(df_sgr$VESSEL_NAME)
max(gurthorm$SALINITY_MEAS)
df_sgr <- read.table("C:\\Workspace\\GIT\\DATA\\Infographic\\df_sgr.csv")
head(df_sgr)
```

```{R}
unique(df_sgr$VESSEL_NAME)
max(gurthorm$SALINITY_MEAS)
unique(df_sgr$VESSEL_NAME)
```


```{R}
# subset to get only gurthorm data
gurthorm <- df_sgr[df_sgr$VESSEL_NAME == 'GUTHORM MAERSK',]
max(gurthorm$SALINITY_MEAS)
min(gurthorm$SALINITY_MEAS)

# view new svg
minisvg::knit_print.SVGDocument(svg_total)
```


```{R}
# max salinity option (this is also the number of bins)
max_sal = 32
#colors <- colorspace::diverging_hcl(n = max_sal, palette = 'Blue-Red')
colors <- colorspace::divergingx_hcl(n = max_sal, palette = 'RdYlBu')

# create a new column for the instrument name that matches the assigned id
gurthorm$INS_NAME <- paste(sub(".*200-", "", gurthorm$STATION_NAME), gurthorm$SALINITY_GAGE_NAME, sep="_")
gurthorm$COLOR <- ""
max_sal <- max_sal - 1

# assign color column based on bins and diverging colors
for(i in 0:max_sal) {
  gurthorm$COLOR[gurthorm$SALINITY_MEAS >= i & gurthorm$SALINITY_MEAS < i + 1] <- colors[i + 1]
}

minisvg::knit_print.SVGDocument(svg_total)
```


```{R}
# create df to compare and QA check
color_df <- data.frame(colors)
color_df$sal_min <- 0:max_sal
color_df$sal_max <- 1:(max_sal + 1)
values <- df_svgId$id
for (value in values) {
df_svgId$count[df_svgId$id == value] <- length(gurthorm$INS_NAME[gurthorm$INS_NAME == value])
}
# select only the rows with the same number of readings
df_svgId_animate_options <- df_svgId[df_svgId$count == 2269,]
# test to see if readings were taken at the same time. count the number of unique readings where instrument id in the options list
print(length(unique(gurthorm$READING_TIME)))

```


```{R}
values <- df_svgId$id
for (value in values) {
df_svgId$count[df_svgId$id == value] <- length(gurthorm$INS_NAME[gurthorm$INS_NAME == value])
}
# select only the rows with the same number of readings
df_svgId_animate_options <- df_svgId[df_svgId$count == 2269,]
# test to see if readings were taken at the same time. count the number of unique readings where instrument id in the options list
print(length(unique(gurthorm$READING_TIME)))
# count unique readings in the dataset
df_unique_readings <- unique(data.frame(gurthorm$READING_TIME))
values <- df_unique_readings$gurthorm.READING_TIME
for (value in values) {
  df_unique_readings$count[df_unique_readings$gurthorm.READING_TIME == value] <- length(gurthorm$INS_NAME[gurthorm$READING_TIME == value])
}
```


```{R}
print(length(unique(gurthorm$INS_NAME)))
print(unique(gurthorm$INS_NAME))
```

```{R}
# import as an svg_doc object
svg_total <- minisvg::parse_svg_doc("man/figures/upper_lock_cross_section.svg")
df_svgId <- get_ids(svg_total)
head(df_svgId)
View(df_svgId)
```


```{R}
df_sgr <- read.table("C:\\Workspace\\GIT\\DATA\\Infographic\\df_sgr.csv")
head(df_sgr)

unique(df_sgr$VESSEL_NAME)
# subset to get only gurthorm data
gurthorm <- df_sgr[df_sgr$VESSEL_NAME == 'GUTHORM MAERSK',]
max(gurthorm$SALINITY_MEAS)
min(gurthorm$SALINITY_MEAS)
```

```{R}
# view new svg
minisvg::knit_print.SVGDocument(svg_total)
```

```{R}

```

```{R}

```


```{R}

```

